---
title: "Heart rate detection and stress"
author: "Tamas Nagy"
date: "`r Sys.Date()`"
output: 
  html_document:
   theme: spacelab
   code_download: true
   toc: true
   toc_float: true
editor_options:
  chunk_output_type: console
---

```{r, echo = TRUE, eval = FALSE, include = FALSE}
# TODO: Separate data processing and analysis

install.packages(c("tidyverse", "skimr", "lme4", "performance", "sjPlot", "here", "emo", "broom.mixed", "psych", "easystats", "ggbeeswarm", "skmir", "lavaan", "tidySEM", "ggExtra"))
```

# Project setup 

```{r setup, message = FALSE, warning = FALSE, include = FALSE}
library(tidyverse)
library(lmerTest)
library(performance)
library(sjPlot)
library(broom.mixed)
library(googlesheets4)
library(stringi)
library(gt)
library(ggbeeswarm)
library(correlation)
library(here)
library(patchwork)
library(skimr)
library(ggsignif)
library(insight)
library(tidyverse)
library(lavaan)
library(tidySEM)
library(semTable)
library(parameters)

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 9, fig.height = 6)

theme_set(theme_light())
```


```{r}
# Set parameters
outlier_treshold_sd = 3 # Threshold for outliers in SD
# Column order for model summaries
col_order <- c("est", 
               # "ci", 
               "se",
               # "df.error", 
               "stat", 
               "p")

renamed_markers <-
  c("baseline" = "Baseline",
    "physical" = "Handgrip",
    "cognitive" = "N-back",
    "mental" = "N-back")

cardioceptive_names <-
  c(ca = "Heartbeat perception",
    hrp = "Heart rate perception")

phys_metrics <- 
  tribble(~name, ~metric,
         "meanhr", "Mean HR (bpm)",
         "lg_hf_percent", "HF% (log %)",
         "lg_rmssd", "RMSSD (log ms)",
         "lg_eda_avg", "Mean EDA (log uS)",
         "hf.percent", "HF% (%)",
         "rmssd", "RMSSD (ms)",
         "eda.avg", "Mean EDA (uS)") %>% 
  mutate(metric = fct_inorder(metric))

# skin_avg <- c("lg_eda_avg", "lg_scl_avg", "lg_scr_avg")
# skin_increase <- c("eda_increase", "scl_increase", "scr_increase")
# heart_metrics <- c("lg_rmssd", "lg_hf_percent", "lg_hf_n_u", "meanhr")


react_vars <-
   c(
      "meanhr_N-back",
      "meanhr_Handgrip",
      "hf.percent_N-back",
      "hf.percent_Handgrip",
      "rmssd_N-back",
      "rmssd_Handgrip",
      "eda.avg_N-back",
      "eda.avg_Handgrip",
      "stai_N-back",
      "stai_Handgrip"
   )

```


```{css, eval = FALSE}
.main-container {
    max-width: 100%;
}
```

The following analysis report contains the descriptive and inferential statistics for the experiment that investigated how physiological and subjective arousal might be mediated by cardioceptive accuracy.

# Reading processed data
```{r}
reactivity <- read_csv(here("data/processed/reactivity_clean.csv"))
experiment <- read_csv(here("data/processed/survey_clean.csv"))
phys_trans <- read_csv(here("data/processed/physiology_clean.csv"))
```

# Sample characteristics

```{r}
experiment %>%
  select(male = sex,
         age,
         stai_s_t0_sum,
         bmi,
         bodyfat_percent,
         max_grip_strength,
         resting_hr) %>%
  mutate(male = if_else(male == "Male", 100, 0),
         n = n(),
         .before = 1) %>%
  skim() %>%
  select( name = skim_variable,
          mean = numeric.mean,
          sd = numeric.sd,
          min = numeric.p0,
          max = numeric.p100,
          Distribution = numeric.hist) %>%
  mutate(name = c("N",
                  "Male %",
                  "Mean age (years)",
                  "Mean STAI-S baseline",
                  "Mean BMI (kg/m<sup>2</sup>)",
                  "Mean bodyfat %",
                  "Mean max grip strength (kg)",
                  "Mean resting HR (bpm)")) %>%
  gt() %>%
  fmt_number(c(mean, sd, min, max), decimals = 1, rows = 2:8) %>%
  fmt_markdown(columns = 1) %>%
  cols_merge(columns = c("mean", "sd"), pattern = "{1} ({2})") %>%
  cols_merge(columns = c("min", "max"), pattern = "{1} - {2}") %>%
  cols_label("name" = "",
             "mean" = "Value (SD)",
             "min" = "Range")

```


# Subjective response to stress

We tested if the tasks were perceived as stressful.  

## STAI-S

Controlling for the baseline level of anxiety (measured in the beginning of the session) and the order of the tasks, we found a significant interaction of task and measurement, i.e. the cognitive task was associated with a STAI-S increase over time, while the physical task did not show a significant change.

```{r}
# Calculate STAI-S reactivity
stai_reactivity <-
   experiment %>% 
   select(id, stai_s_mental_t0_sum,  stai_s_mental_t1_sum, stai_s_physical_t0_sum, stai_s_physical_t1_sum) %>% 
   transmute(id,
             `stai_react_N-back` = stai_s_mental_t1_sum - stai_s_mental_t0_sum,
             stai_react_Handgrip = stai_s_physical_t1_sum - stai_s_physical_t0_sum)

   
stai <-
   experiment %>% 
   select(id, ends_with("_sum"), task_order) %>% 
   pivot_longer(cols = -c(id, stai_s_t0_sum, task_order),
                names_to = c("task", "time"),
                names_pattern = "stai_s_(.*)_(.*)_sum") %>% 
   mutate(task = recode(task, !!!renamed_markers))

stai_sum <-
   stai %>% 
   group_by(task, time) %>% 
   summarise(stai_avg = mean(value, na.rm = TRUE),
             stai_sd = sd(value, na.rm = TRUE),
             stai_se = stai_sd/sqrt(n())) %>% 
   mutate(stai_prop = stai_avg/first(stai_avg)) %>% 
   ungroup()

stai_sum %>% 
   select(-stai_se) %>% 
   pivot_wider(names_from = c(time), names_sep = "_",
               values_from = c(stai_avg, stai_sd, stai_prop)) %>% 
   select(-stai_prop_t0) %>% 
   gt() %>% 
   fmt_number(columns = starts_with("stai_"), 
                 decimals = 1) %>% 
   fmt_percent(stai_prop_t1, decimals = 1) %>% 
   cols_merge(columns = ends_with("_t0"), 
              pattern = "{1} ± {2}") %>% 
   cols_merge(columns = ends_with("_t1"), 
              pattern = "{1} ± {2} ({3})") %>% 
   cols_label(task = "Task",
              stai_avg_t0 = "Pre-stress", 
              stai_avg_t1 = "Post-stress") %>% 
   tab_options(column_labels.font.weight = "bold",
               column_labels.background.color = "lightgrey") %>%
   tab_footnote(footnote = "Values represent the mean (SD) of the STAI-S",
                locations = cells_column_labels(3))

```

<div align="center">
```{r}
# Stats
stai %>% 
   filter(task != "baseline") %>% 
   lmer(value ~ task * time + stai_s_t0_sum + task_order + (1|id), 
        data = .) %>% 
   tab_model(dv.labels = "STAI-S", 
             show.aicc = TRUE,
             show.stat = TRUE,
             show.df = TRUE,
             show.std = TRUE)
```
</div>

```{r}
# Visualize
stai_sum %>% 
   mutate(time = recode(time, t0 = "Baseline", t1 = "Task")) %>% 
   ggplot() +
   aes(x = time, 
       y = stai_avg, 
       group = task, 
       color = task,
       ymin = stai_avg - stai_se, 
       ymax = stai_avg + stai_se) +
   geom_errorbar(width = .1, alpha = .7, color = "gray50") +
   geom_point() +
   geom_line(size = 1.2, alpha = .6) +
   # scale_y_continuous(limits = c(0, NA)) +
   labs(color = "Task",
        # title = "Change in STAI-S by task and time",
        # subtitle = "STAI-S increased only in the cognitive stress task.",
        y = "Mean STAI-S (SEM)",
        x = NULL) +
  theme(panel.grid = element_blank())


```

```{r}
metric_order <- c("STAI-S Before", "STAI-S After", "Expected stress", "Expected HR", 
                  "Perceived HR")

stai_sum_wide <-
  experiment %>% 
  select(id, 
         # I just added _xy_ because it is later easier to match the pattern this way
         `STAI-S Before_xy_N-back` = stai_s_mental_t0_sum,  
         `STAI-S After_xy_N-back` = stai_s_mental_t1_sum, 
         `STAI-S Before_xy_Handgrip` = stai_s_physical_t0_sum, 
         `STAI-S After_xy_Handgrip` = stai_s_physical_t1_sum)

subj_wide <-
  experiment %>% 
  select(id, 
         `Expected stress_change_N-back` = mental_expected_stress,
         `Expected stress_change_Handgrip`= physical_expected_stress,
         # `Perceived stress_change_N-back` = mental_experienced_stress,
         # `Perceived stress_change_Handgrip` = physical_experienced_stress,
         `Expected HR_change_N-back` = mental_expected_hr_change,
         `Expected HR_change_Handgrip` = physical_expected_hr_change,
         `Perceived HR_change_N-back` = mental_experienced_hr_change,
         `Perceived HR_change_Handgrip` = physical_experienced_hr_change) %>% 
  left_join(stai_sum_wide, by = "id")

subj_long <-
  subj_wide %>% 
  pivot_longer(cols = -c("id"),
               names_to = c("metric", NA,"task"),
               names_pattern = "^(.*)_(.*)_(.*)",
               values_to = "value")

# Correlations of subj metrics
handgrip_cor <-
  subj_wide %>% 
  select(ends_with("_Handgrip")) %>%
  rename_with(~str_remove(., "_change|_xy")) %>% 
  correlation(redundant = TRUE) %>% 
  summary(redundant = TRUE) %>% 
  as_tibble() %>% 
  separate(Parameter, into = c("metric", "task"), sep = "_") %>% 
  rename_with(~str_remove(., "_Handgrip"))

nback_cor <-
  subj_wide %>% 
  select(ends_with("_N-Back")) %>%
  rename_with(~str_remove(., "_change|_xy")) %>% 
  correlation(redundant = TRUE) %>% 
  summary(redundant = TRUE) %>% 
  as_tibble() %>% 
  separate(Parameter, into = c("metric", "task"), sep = "_") %>% 
  rename_with(~str_remove(., "_N-back"))

all_cor <-
  bind_rows(handgrip_cor, nback_cor)

subj_long %>% 
  group_by(task, metric) %>% 
  summarise(Mean = mean(value, na.rm = TRUE),
            SD = sd(value, na.rm = TRUE), .groups = "drop") %>% 
  left_join(all_cor, by = c("task", "metric")) %>% 
  mutate(metric = fct_relevel(metric, !!!metric_order)) %>% 
  arrange(task, metric) %>% 
  select(task, metric, Mean, SD, all_of(metric_order)) %>% 
  gt() %>% 
  fmt_number(c("Mean", "SD"), decimals = 1) %>% 
  fmt_number(`STAI-S Before`:`Perceived HR`, decimals = 2)
```

The subjective stress measure also showed, that participants regarded the mental stressor to be more stressful than the physical stressor.

```{r}
experiment %>% 
   select(id, mental_experienced_stress, physical_experienced_stress, stai_s_t0_sum, task_order) %>% 
   pivot_longer(-c(id, stai_s_t0_sum, task_order)) %>% 
   mutate(name = str_remove(name, "_experienced_stress") %>% 
                 recode(!!!renamed_markers)) %>% 
  lmer(value ~ name + stai_s_t0_sum + task_order + (1|id), data = .) %>% 
     tab_model(dv.labels = "STAI-S", 
             show.stat = TRUE,
             show.df = TRUE,
             show.std = TRUE)

experiment %>% 
   select(id, mental_experienced_stress, physical_experienced_stress) %>% 
   pivot_longer(-id) %>% 
   mutate(name = str_remove(name, "_experienced_stress") %>% 
                 recode(!!!renamed_markers)) %>% 
  t.test(value ~ name, data = ., paired = TRUE)

```

# Psychophysiological response to stress

```{r}
phys_trans %>% 
   select(starts_with("lg_")) %>% 
   pivot_longer(everything()) %>% 
   ggplot() +
   aes(x = value) +
   geom_histogram() +
   facet_wrap(~name, scales = "free") +
   labs(x = NULL, y = NULL,
        title = "Histogram of the log+1 transformed physiological variables",
        subtitle = "Outliers have been removed")


```

```{r}
reactivity %>% 
   pivot_longer(-id) %>% 
   ggplot() +
   aes(x = value) +
   geom_histogram() +
   facet_wrap(~name, scales = "free", nrow = 4) +
   labs(title = "Reactivity histograms of key metrics")

```

# Analysis of the effects of the tasks on physiological variables
## Descriptive statistics
```{r}

phys_trans %>% 
  select(id, marker, ca, hrp, 
         meanhr, lg_hf_percent, lg_rmssd, lg_eda_avg) %>% 
  summarise(across(ca:lg_eda_avg, 
                   list(mean = mean, sd = sd), 
                   na.rm = TRUE)) %>% 
  pivot_longer(everything(), 
               names_to = c("name", "stat"), 
               names_pattern = "^(.*)_(.*)$") %>% 
  pivot_wider(names_from = "stat") %>% 
  left_join(phys_metrics, by = "name") %>% 
  relocate(metric) %>% 
  gt() %>% 
  fmt_number(mean:sd, decimals = 2)
```



## Visualize phys changes
```{r}
# Create table for annotation
annotation_df <- 
  phys_trans %>% 
  select(id, marker, meanhr, lg_hf_percent, lg_rmssd, lg_eda_avg) %>% 
  pivot_longer(meanhr:lg_eda_avg, values_drop_na = TRUE) %>%
  left_join(phys_metrics, by = "name") %>% 
  group_by(metric) %>%
  nest() %>% 
  summarise(model = map(data, ~lmer(value ~ marker + (1|id), data = .x) %>% 
                               broom.mixed::tidy())) %>% 
  unnest(model) %>% 
  filter(effect == "fixed", term != "(Intercept)") %>% 
  mutate(term = str_remove(term, "marker")) %>% 
  transmute(metric, 
            start = "Baseline",
            end = term, 
            marker = end,
            label = paste0("p=",scales::pvalue(p.value)) %>% 
                    str_replace("=<", "<"),
            y = c(81.7, 82.4, 3.17, 3.21, 3.64, 3.665, 2.12, 2.16)) %>% 
  group_by(metric) %>% 
  mutate(top = max(y) * 1.015) %>% 
  ungroup()


# Plot physiological changes with significance
phys_trans %>% 
  select(id, marker, meanhr, lg_hf_percent, lg_rmssd, lg_eda_avg) %>% 
  pivot_longer(meanhr:lg_eda_avg, values_drop_na = TRUE) %>%
  left_join(phys_metrics, by = "name") %>% 
  ggplot() +
  aes(x = marker, y = value, group = metric, color = metric) +
  stat_summary(fun.data = "mean_se", show.legend = FALSE, 
               geom = "errorbar", color = "gray", width = .2) +
  stat_summary(fun.data = "mean_se", geom = "point", show.legend = FALSE) +
  stat_summary(fun.data = "mean_se", show.legend = FALSE, geom = "line", 
               size = 1.2, alpha = .6) +
  facet_wrap(~metric, scales = "free_y") +
  labs(y = "Mean value (SEM)",
       x = NULL) +
  theme(panel.grid = element_blank()) +
  geom_blank(data = annotation_df, aes(y = top)) +
  geom_signif(inherit.aes = FALSE, 
              data = annotation_df, manual = TRUE, 
              aes(xmin = start, xmax = end, annotations = label, y_position = y),
              color = "black", alpha = .6,
              textsize = 3, tip_length = .002, margin_top = 12, vjust = -.1)
```

## Change  stats in %
```{r}

phys_trans %>% 
  select(id, marker, one_of(phys_metrics$name[1:4])) %>% 
  pivot_longer(-c(id, marker)) %>% 
  group_by(name, marker) %>% 
  summarise(avg = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>% 
  mutate(prop_change = avg / first(avg) %>% round(2)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = marker,
              values_from = c(avg, sd, prop_change)) %>% 
  left_join(phys_metrics, by = "name") %>% 
  relocate(metric) %>% 
  select(-name) %>% 
  arrange(metric) %>% 
  gt() %>% 
  fmt_percent(starts_with("prop_"), decimals = 1) %>% 
  fmt_number(matches("^avg.*|^sd.*"), decimals = 2) %>%
  cols_merge(columns = ends_with("_Baseline"),
             pattern = "{1} ± {2}") %>% 
  cols_merge(columns = ends_with("_Handgrip"),
             pattern = "{1} ± {2} ({3})") %>% 
  cols_merge(columns = ends_with("_N-Back"),
             pattern = "{1} ± {2} ({3})") %>% 
  cols_label(metric = "",
             avg_Baseline = "Baseline",
             avg_Handgrip = "Handgrip",
             `avg_N-back` = "N-back") %>% 
  cols_align(metric, align = "left")
  
```

## Reactivity measures
### Variables
Reactivity (task - baseline) values were calculated for EDA average, HF%, RMSSD, HR average, and STAI. 

### Results

```{r}
reactivity %>% 
   pivot_longer(-id, values_drop_na = TRUE) %>% 
   group_by(name) %>% 
   summarise(avg = mean(value),
             se = sd(value)/sqrt(n())) %>% 
   separate(name, 
            into = c("metric", "marker"), 
            sep = "_",
            extra = "merge") %>% 
   mutate(marker = str_remove_all(marker, "react_")) %>% 
   ggplot() +
   aes(x = marker, y = avg, ymin = avg - se, ymax = avg + se, 
       group = metric) +
   geom_pointrange() +
   geom_line(aes(color = metric), size = 1.2, alpha = .6, 
             show.legend = FALSE) +
   geom_hline(yintercept = 0, color = "red", lty = "dashed") +
   facet_wrap(~metric, scales = "free") +
   labs(title = "Reactivity measures: Difference from baseline by task",
        subtitle = "Zero (dashed red line) shows baseline",
        x = NULL,
        y = "Mean (SD)")

```

# Creating tables and visualizations


```{r}
experiment %>% 
   select(id, ca, hrp) %>% 
   left_join(reactivity, by = "id") %>% 
   rename_with(~str_remove(., "react_")) %>% 
   correlation(p_adjust = "fdr", 
               select2 = c("ca", "hrp"),
               select = all_of(react_vars)) %>% 
   summary() %>% 
   separate(Parameter, into = c("name", "task"), sep = "_") %>% 
   left_join(phys_metrics, by = "name") %>% 
   filter(!str_detect(name, "stai")) %>% 
   relocate(task, metric) %>% 
   select(-name) %>% 
   arrange(task) %>% 
   as_tibble() %>% 
   print_html() %>% 
   cols_hide(1)
```

## HBP and HRP corr figure
```{r}

hrp_hbp <-
  experiment %>% 
  select(hrp, ca) %>% 
  ggplot() +
  aes(x = ca, y = hrp) +
  geom_point(alpha = .8) +
  geom_smooth(method = lm) +
  labs(x = "Heart beat perception",
       y = "Heart rate perception") +
  theme(panel.grid = element_blank()) +
  coord_fixed(ratio = 1)
  

ggExtra::ggMarginal(hrp_hbp, type = "density", alpha = .7,
                    xparams = list(fill = "salmon"),
                    yparams = list(fill = "cyan")
                    )
```

## Correlation between pooled physiological measurements and cardioceptive metrics

```{r eval=FALSE, include=FALSE}

experiment %>% 
   select(id, ca, hrp) %>% 
   left_join(select(phys_trans, id, marker, one_of(phys_metrics$name[1:4])), 
             by = "id") %>% 
   filter(marker == "Baseline") %>% 
   correlation(p_adjust = "fdr", 
               select2 = c("ca", "hrp"),
               select = c("meanhr","lg_hf_percent","lg_rmssd","lg_eda_avg")) %>% 
   left_join(phys_metrics, by = c("Parameter1" = "name")) %>% 
   relocate(metric, .after = Parameter2) %>% 
   select(-Parameter1) %>%
   arrange(Parameter2) %>%
   as_tibble() %>% 
   print_html()
  
# Simplified table
experiment %>% 
   select(id, ca, hrp) %>% 
   left_join(select(phys_trans, id, marker, one_of(phys_metrics$name[1:4])), 
             by = "id") %>% 
   mutate(marker = factor(marker),
          id = factor(id)) %>% 
   correlation(p_adjust = "fdr", 
               select = c(#"id",
                          "ca", "hrp",
                          "meanhr","lg_hf_percent","lg_rmssd","lg_eda_avg"),
               # include_factors = TRUE,
               # multilevel = TRUE
               ) %>% 
   summary() %>% 
   left_join(phys_metrics, by = c("Parameter" = "name")) %>% 
   as_tibble() %>% 
   print_html() %>% 
   cols_move_to_start(metric) %>% 
   cols_hide("Parameter") %>% 
   cols_label(metric = "Measurement",
              meanhr = "Mean HR",
              lg_hf_percent = "HF% (log)",
              lg_rmssd = "RMSSD (log)",
              lg_eda_avg = "Mean EDA (log)",
              hrp = "HRP")

```


We did not find a relationship between cardioceptive accuracy and physiological reactivity measures (HR, RMSSD, HF%, EDA) in any of the tasks. However, CA was negatively associated with HR and positively with lgRMSSD.
Similarly, heart rate perception was negatively associated with HR and positively with lgRMSSD, but did not moderate the reaction to stress.

```{r}
phys_trans %>% 
   select(meanhr, lg_rmssd, marker, hrp, ca) %>% 
   pivot_longer(meanhr:lg_rmssd, 
                names_to = "name",
                values_to = "value") %>% 
   pivot_longer(hrp:ca, 
                names_to = "cardio_metric", 
                values_to = "cardio_value") %>% 
   left_join(phys_metrics, by = "name") %>% 
   mutate(cardio_metric = recode(cardio_metric, !!!cardioceptive_names)) %>% 
   ggplot() +
   aes(x = cardio_value, y = value, color = marker) +
   geom_point(alpha = .6) +
   geom_smooth(method = "lm", 
               size = 1.2, 
               se = FALSE, 
               alpha = .6) +
   scale_color_viridis_d() +
   labs(color = "Task",
        y = NULL,
        x = NULL) +
   facet_grid(metric ~ cardio_metric, scales = "free") +
   theme(panel.grid = element_blank())
```


# Hypothesis tests
## H1-2
Perceived stress in both conditions [mental_experienced_stress; physical_experienced_stress / STAI-S_mental_after – STAI-S_mental_before; STAI-S_physical_after - STAI-S_physical_before] is predicted by cardioceptive accuracy [cardioceptive_accuracy], physiological reactions [ΔHRmental, ΔEDAmental, ΔRMSSDmental, ΔHFmental; ΔHRphysical, ΔEDAphysical, ΔRMSSDphysical, ΔHFphysical], and stress related expectations [mental_expected_stress; physical_expected_stress]

We fitted linear regressions for each task (N-back and physical) separately. We investigated if physiological changes, cardioceptive accuracy (HBP), or expected stress predicted STAI-S change in the respective tasks. We used task order as a covariate. 

- HRP was negatively associated with STAI-S change (in the Handgrip task).
- Expected stress predicted STAI-S change in the N-back task (but not the Handgrip task).
- From the physiological measures, only RMSSD predicted STAI-S change (in the N-back task).
- The interaction of HRP and mean HR negatively predicted STAI-S change (in the Handgrip task).
- The interaction of HRP and mean HF% negatively predicted STAI-S change (in the N-back task).


```{r}
hypo12 <-
   reactivity %>% 
   left_join(select(experiment, 
                    id, ca, hrp, 
                    bodyfat_percent, 
                    sex, task_order,
                    physical_expected_stress,
                    mental_expected_stress),
                    by = "id") %>%
   pivot_longer(`meanhr_react_N-back`:eda.avg_react_Handgrip,
                names_pattern = "(.*)_react_(.*)",
                names_to = c("metric", "task")) %>% 
   group_by(metric, task) %>% 
   # Scale all numeric vaiarables (by metric and task)
   mutate(across(c(`stai_react_N-back`:bodyfat_percent,                  
                 physical_expected_stress:mental_expected_stress,
                 value), 
                 ~scale(.x) %>% 
                  as.numeric())) %>%    
   nest() %>% 
   mutate(`N-back_stai_react_ca` = map(data, 
                                     ~lm(`stai_react_N-back` ~ ca * value +
                                         mental_expected_stress + 
                                         bodyfat_percent +
                                         sex + 
                                         task_order,
                                         data = .x)),
          Handgrip_stai_react_ca = map(data, 
                                     ~lm(stai_react_Handgrip ~ ca * value +
                                         physical_expected_stress + 
                                         bodyfat_percent +
                                         sex + 
                                         task_order,
                                         data = .x)),
          `N-back_stai_react_hrp` = map(data, 
                                     ~lm(`stai_react_N-back` ~ hrp * value +
                                         mental_expected_stress + 
                                         bodyfat_percent +
                                         sex + 
                                         task_order,
                                         data = .x)),
          Handgrip_stai_react_hrp = map(data, 
                                     ~lm(stai_react_Handgrip ~ hrp * value + 
                                         physical_expected_stress + 
                                         bodyfat_percent +
                                         sex + 
                                         task_order,
                                         data = .x))) %>% 
   pivot_longer(`N-back_stai_react_ca`:Handgrip_stai_react_hrp) %>% 
   filter((str_detect(task, "N-back") & str_detect(name, "N-back")) |
          (str_detect(task, "Handgrip") & str_detect(name, "Handgrip"))) %>% 
   mutate(name = str_replace(name, "N-back", task)) %>% 
   mutate(name = str_replace(name, "react", metric)) %>% 
   ungroup()
   
hypo12_sum <- 
   hypo12 %>% 
   pull(value) %>% 
   set_names(hypo12$name) %>% 
   map_dfr(tidy, conf.int = TRUE, .id = "model") %>% 
   group_by(model) %>% 
   mutate(metric = str_match(model, ".*_(.*)_.*$")[,2],
          term = str_replace(term, "value", metric)) %>% 
   select(model, term, estimate, std.error, conf.low, conf.high, statistic, p.value) 

# All models  (to Supplementary material)
hypo12_sum %>% 
   gt() %>% 
   fmt_number(c("estimate", "std.error", "conf.low", "conf.high"), decimals = 2) %>%
   fmt_number(c("statistic", "p.value"), decimals = 3) %>%
   tab_options(column_labels.font.weight = "bold",
               column_labels.background.color = "lightgrey",
               row_group.background.color = "lightgrey", 
               row_group.font.weight = "bold", 
               row.striping.include_table_body = TRUE,
               row.striping.background_color = "#EEEEEE"
   )   

hypo12 %>% 
  mutate(model_stat = map(value, glance)) %>% 
  unnest(model_stat)

### All significant associations

hypo12_sum %>% 
  ungroup() %>% 
  # mutate(p.value = p.adjust(p.value, method = "fdr")) %>%
  filter(p.value < .05,
         !term %in% c("(Intercept)", "bodyfat_percent", "sexMale")) %>%
  arrange(model, term) %>% 
     gt() %>% 
   fmt_number(c("estimate", "std.error", "conf.low", "conf.high"), decimals = 2) %>%
   fmt_number(c("statistic", "p.value"), decimals = 3) %>%
   tab_options(column_labels.font.weight = "bold",
               column_labels.background.color = "lightgrey",
               row_group.background.color = "lightgrey", 
               row_group.font.weight = "bold", 
               row.striping.include_table_body = TRUE,
               row.striping.background_color = "#EEEEEE"
   )   

### Simplified table

hypo12_sum %>% 
  separate(model, into = c("task", NA, "name", "ca"), sep = "_") %>%
  left_join(phys_metrics, by = "name") %>% 
  filter(!term %in% c("(Intercept)", "ca", "hrp", "task_orderPhysical first", "sexMale","bodyfat_percent")) %>% 
  mutate(term = str_remove(term, "mental_|physical_") %>% 
                str_replace("hrp", "ca") %>% 
                str_replace("meanhr|hf.percent|rmssd|eda.avg", "phys_change"),
         sig = case_when(p.value <= .001 ~ "***",
                         p.value <= .01 ~ "**",
                         p.value <= .05 ~ "*",
                         TRUE ~ "")) %>% 
  transmute(task, ca, metric, term, 
            estimate, sig
            ) %>% 
  pivot_wider(names_from = term, values_from = c(estimate, sig)) %>% 
  arrange(task, ca, metric) %>% 
  gt(rowname_col = "ca",
     groupname_col = "task") %>% 
  cols_align(columns = metric, align = "left") %>% 
  fmt_number(columns = c(estimate_phys_change, estimate_expected_stress, `estimate_ca:phys_change`), decimals = 2) %>% 
  cols_merge(columns = c(estimate_phys_change, sig_phys_change)) %>% 
  cols_merge(columns = c(estimate_expected_stress, sig_expected_stress)) %>% 
  cols_merge(columns = c(`estimate_ca:phys_change`, `sig_ca:phys_change`))

```


### Figure: CA/HRP moderates the association of objective and subjective change metrics

```{r}
interaction_df <- 
  reactivity %>% 
  select(id, `meanhr_react_N-back`, meanhr_react_Handgrip,
         `stai_react_N-back`, stai_react_Handgrip) %>%
  left_join(select(experiment, id, ca, hrp, hrp_group, hbp_group), by = "id") %>%
  pivot_longer(`meanhr_react_N-back`:stai_react_Handgrip,
               names_to = c("name", "task"),
               names_pattern = "(.*_react)_(.*)") %>% 
  pivot_wider(names_from = "name", values_from = "value")

hbp_plot <- 
  interaction_df %>% 
  ggplot() +
  aes(y = stai_react, x = meanhr_react, color = hbp_group) +
  geom_hline(yintercept = 0, lty = "dashed", alpha = .2) +
  geom_vline(xintercept = 0, lty = "dashed", alpha = .2) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm", se = FALSE, alpha = .7) +
  scale_color_brewer(palette = "Reds") +
  facet_wrap(~task, scales = "free_x") +
  labs(color = "HBP group",
       x = NULL,
       y = "STAI change") +
  theme(legend.position = c(0.93, 0.22), panel.grid = element_blank())

hrp_plot <-
  interaction_df %>% 
  ggplot() +
  aes(y = stai_react, x = meanhr_react, color = hrp_group) +
  geom_hline(yintercept = 0, lty = "dashed", alpha = .2) +
  geom_vline(xintercept = 0, lty = "dashed", alpha = .2) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm", se = FALSE, alpha = .7) +
  scale_color_brewer(palette = "Blues") +
  facet_wrap(~task, scales = "free_x") +
  labs(color = "HRP group",
       x = "HR change",
       y = "STAI change") +
  theme(legend.position = c(0.93, 0.22), panel.grid = element_blank())

hbp_plot + hrp_plot + plot_layout(nrow = 2)
  
```

## H3 
Perceived HR in both conditions [mental_experienced_HR_change; physical_experienced_HR_change] is predicted by actual change [ΔHRmental, ΔHRphysical] and HR-related expectations [mental_expected_HR_change; physical_expected_HR_change] in both conditions

We used a path analysis to include all variables in the same model. We included sex, body fat percent, and task order as covariates.

We found that actual HR change did not predict the experienced HR change in any of the tasks. However, expected HR change significantly predicted subjective HR change in both tasks. STAI was only predicted for the N-Back a


```{r}
# Prepare data
path_df <-
   reactivity %>% 
   select(id, starts_with("meanhr_"),
          starts_with("rmssd_"),
          starts_with("hf.percent_"),
          starts_with("eda.avg_"),
          stai_react_Handgrip,
          `stai_react_N-back`) %>% 
   # Put HR reactivity metrics expected and experienced change together
   left_join(select(experiment, 
                    id, bodyfat_percent, sex, task_order, hrp, ca,
                    `expectedhr_change_Handgrip` = physical_expected_hr_change,
                    `experiencedhr_change_Handgrip` = physical_experienced_hr_change,
                    `expectedhr_change_N-back` = mental_expected_hr_change,
                    `experiencedhr_change_N-back` = mental_experienced_hr_change), 
             by = "id") %>%
   # Create a redundant dataset to run several regressions at once
   pivot_longer(cols = -c("id", "bodyfat_percent", "sex", "task_order", "hrp", "ca"),
                names_to = c("metric","task"),
                names_pattern = "^(.*)_.*_(.*)",
                values_to = "value") %>%  
  pivot_wider(names_from = "metric", 
              values_from = "value") %>% 
  mutate(across(where(is.numeric), ~scale(.x, scale = FALSE)[,1]))

hr_model <- 
  "stai ~ expectedhr + experiencedhr + meanhr + task_order
   experiencedhr ~ expectedhr + meanhr + task_order
   expectedhr ~ task_order
   meanhr ~ expectedhr + task_order + sex + bodyfat_percent
"

# Fit model in lavaan
hr_path <- sem(model = hr_model, 
               data = path_df, 
               group = "task")

# Return results
summary(hr_path, standardized = TRUE, fit.measures = TRUE)
# parameterEstimates(hr_path, standardized = TRUE, rsquare = TRUE) %>% 
#   write_csv(file = here::here("docs/path_models/h3.csv"))

parameters(hr_path, standardize = "std.all") %>% print_html() 

# Plotting SEM
# Format node labels
pretty_nodes <-
  get_nodes(hr_path, label = "name") %>% 
  mutate(label = recode(label,
                        expectedhr = "Expected\nHR change",
                        experiencedhr = "Perceived\nHR change",
                        stai = "ΔSTAI",
                        meanhr = "ΔHR")) %>% 
  # Remove confounders
  filter(!str_detect(name, "task_order|sex|bodyfat_percent"))

pretty_edges <- 
                  # Format edges (to std)
  get_edges(hr_path, label = "est_sig_std") %>% 
                  # Remove variances
                  filter(op != "~~",
                  # Remove confounders
                         !str_detect(lhs, "task_order|sex|bodyfat_percent"),
                         !str_detect(rhs, "task_order|sex|bodyfat_percent")
                  )
  

# Create a layout
layout <- get_layout("", "", "", "stai", "", "","", 
                     "", "", "", "meanhr", "", "","", 
                     "expectedhr", "", "", "", "", "", "experiencedhr",
                     rows = 3)

# Plotting SEM
hr_plot <- 
  graph_sem(hr_path, 
            edges = pretty_edges,
            nodes = pretty_nodes,
            layout = layout
            )

hr_plot
```




